Discussion Draft

### 2020-04-01 doc/Discussion.rmd
added spindle points

```{r reversed.HetC, echo=FALSE}

@@

```{r within.mouse.var, include=FALSE}

@@

#<spindle model -- the telomeric placement of crossovers, minimizes the amount of sister cohesion connecting homologs>

@@

>>>>

```{r reversed.HetC, echo=FALSE}
1c. sex differences in the SAC strength connected to conserved features of gametogenesis
- Sperm SAC strong (can detect 1 unstable bivalent) and oocyte SAC weak / 'leaky' achiasmate chromosomes pass through and higher rate of annueploidy
- leaky oocyte SAC linked to acentrosome spindle (mulitple aMTOCs) and cell volumne, these are features which differ in spermatocyttes (small, centrosome spindle, 2MTOCs)

1d. directional selection could be seen under strict SAC if the fitness landscape for 'chromosome structure' is altered via genetic architecture changes or GxE effects.
```

@@

```{r within.mouse.var, include=FALSE}
2. Greater between cell variation in females, leaky SAC in oocytes can cause more between cell variation within an individual.
2a. relaxed selection on chromosome structure in oocyte / leaky SAC, can lead to larger variance in crossover number between cells. The opposite dynamics of the purifying selection for 'strict' SAC in spermatocytes.

?2b. (leaky SAC could be adaptive/prevent to selfish genetic elements (Lampson))?

@@

#<spindle model -- the telomeric placement of crossovers, minimizes the amount of sister cohesion connecting homologs>
3. Sexual dimorphism in broad scale recombination landscape,  telomere bias vs uniform recombination landscapes fit with purifying selection and relaxed selection respectively.
3a. There are sex differences in demand for chromosome cohesion at MI, the trait / variable causing variation in fitness landscape across chromosome structures might be the amount of sister-cohesion connecting homologs.  In oocytes the cell cycle is slower and arrested in some species, this likely selects for chromosome structures that have stability on the spindle. (required to start with more since the cohesion isn't replenished. The optimal chromosome structure might have more sister cohesion, however given the leaky SAC, selection may not be effective.

3c. Logic of number and placement modulating this area of sister cohesion and chromosome structure (I plan to make a figure to illustrate)
-For single crossover acrocentric chromosome the sister cohesion distal to the crossover connects homologs. This area reduced when crossover position shifted to telomere (each end of chromosomes). Two crossover acrocentric chromosome the area is distal from the first until the second crossover.

3d. Metaphase to anaphase transition proceeds by proteolytic decay of sister cohesion on 'arms'.
- The telomere bias could have an effect of synchronizing the separation of all the bivalents within a cell.

*caveats: Type II crossovers not accounted for and chromatid interference effects on (multi-crossover chromsomes) could alter the sister cohesion pattern.

@@

5. Interference evolution shows a positive correlation with gwRR
5a. Stronger interference for two crossover chromosomes, increases the amount of sister-cohesion connecting the homologs.
5b. Under spindle selection model, the fitness landscape of optimal chromosome structure may have changed in the high recombining strains resulting in more 2CO, with larger IFD (sister cohesion area).
